\textls[-15]{\thetitle \mbox{(Technical Report)}}

### 2022-06-26 standardmodel.tex
Finishing Language Design Section. Try to fix graph embedding and format.
Save syn-translate
Fix syn-translation and others.

Once we have the syntax of our theory, the first question is if our theory is consistent, i.e., if we can syntactically derive bottom in our theory. We prove the consistency by extending the standard model from~\cite{altkap2016,kaposi2017type, kaposi2019gluing} and their formalization, where they formally resort to the concept of \textit{algebra} of QIIT, here we only consider inductively interpreting each syntax piece into naive set theory that is also respecting judgemental equalities\footnote{In our appendix, we still interpret into Agda component to be more precise and less error-prone, also we need the inductive facility from Agda.} The main idea of the proof is that: 1. we use $\denotesS{}$ to map each intrinsic well-typed syntax piece into a mathematical object; 2. we define proof-irrelevant proposition $\models_S$ for each syntactical $\vdash$, and derive compatibility lemma. We will mainly focus on the definition of $\denotesS{}$ and omit the verification of $\models_S$\footnote{Actually our 1 is acting exactly as an interpretation function, and in our agda formulation we only do 1 because 2 is automatic; here we make 2 explicit to make our formulation more accessible and uniform}, and also we omit the verification of equality (i.e. the judgementally-equal syntax needs to be mapped to the same data for well-definedness).

@@

  \denotesS{\goodType{\Gamma}{\cC \sigma}{}} &= \denotesS{\Sigma} \denotesS{\goodType{\Gamma}{\cC \tau}{}} \denotesS{T[sf \ f]} \\

@@

  &= \denotesS{\goodTerm{\Gamma}{\cCt (p_1\mu \ o)}{...}}, \denotesS{(p_2\mu \ o)[sf \ f][(id, \cCt (p_1\mu \ o))]} \\

@@

We encourage the reader to read the (fake) Agda-style proof in our appendix because most of the parts are similar but with more type annotation so it is clearer about the mathematical object doing the interpretation. Here, we need to clarify some ambiguity in our conventional formulation. For example, when we wrote ``$\denotesS{\goodType{\Gamma}{\cC \sigma}{}}$ defined inductively on $\denotesS{\Sigma}$'', the best way to understand is to consider we are defining a function $\denotesS{\goodType{\Gamma}{\cC \_}{}}$ as the interpretation for $\goodType{\Gamma}{\cC \_}{}$ that mapping $\denotesS{\sigma}$ to $\denotesS{\goodType{\Gamma}{\cC \sigma}{}}$ also defined inductively on the length of the signature. The case is similar for $\denotesS{\cL \_}$, $\denotesS{\cCt \_}$ and  $\denotesS{\Sigma \ \_ \ \_}$. \textbf{This is actually the formulation in Agda}---QIIT syntax consider $\cC : "Sig"~\Gamma~n \to "Ty"~\Gamma$ as a constructor and our model needs to interpret this constructor to a function $\cC^S : "Sig"^S \Gamma^S~n \to "Ty"^S~\Gamma^S$ with each judgement type also interpreted.

@@

\begin{figure}
  \centering
\begin{minipage}{0.8\linewidth}
  \begin{minted}[mathescape, escapeinside=||]{Agda}
    data MW {i}{j} (σ : |$\denotesS{\texttt{WSig i }\cdot\texttt{ S}}$|) : Set (lsuc i) where
      MWsup : |$n$| < S → (a : |$\pi^n_1$| σ) → ((|$\pi^n_2$| σ) a → MW σ) 
              → MW σ
      \end{minted}
\end{minipage}
\caption{Modeling Inductive Type}\label{fig:model-ind-type}
\end{figure}

>>>>

Once we have the syntax of our theory, the first question is if our theory is consistent, i.e., if we can syntactically derive bottom in our theory. We prove the consistency by extending the standard model from~\cite{altkap2016,kaposi2017type, kaposi2019gluing} and their formalization, where they formally resort to the concept of \textit{algebra/model} of QIIT, here we only consider inductively interpreting each syntax piece into naive set theory that is also respecting judgemental equalities\footnote{In our appendix, we still interpret into Agda component to be more precise and less error-prone, also we need the inductive facility from Agda.} The main idea of the proof is that: 1. we use $\denotesS{}$ to map each intrinsic well-typed syntax piece into a mathematical object; 2. we define proof-irrelevant proposition $\models_S$ for each syntactical $\vdash$, and derive compatibility lemma. We will mainly focus on the definition of $\denotesS{}$ and omit the verification of $\models_S$\footnote{Actually our 1 is acting exactly as an interpretation function, and in our agda formulation we only do 1 because 2 is automatic; here we make 2 explicit to make our formulation more accessible and uniform}, and also we omit the verification of equality (i.e. the judgementally-equal syntax needs to be mapped to the same data for well-definedness).

@@

  \denotesS{\goodType{\Gamma}{\cC \sigma}{}} &= \denotesS{\Sigma} \denotesS{\goodType{\Gamma}{\cC \tau}{}} \denotesS{T["sf" \ f]} \\

@@

  &= \denotesS{\goodTerm{\Gamma}{\cCt (p_1\mu \ o)}{...}}, \denotesS{(p_2\mu \ o)["sf" \ f][(id, \cCt (p_1\mu \ o))]} \\

@@

We again encourage the reader to read the (pseudo-)Agda-style proof mentioned in our appendix because most of the parts are similar but with more type annotation so it is clearer about the mathematical object doing the interpretation. Here, we need to clarify some ambiguity in our conventional formulation. For example, when we wrote ``$\denotesS{\goodType{\Gamma}{\cC \sigma}{}}$ defined inductively on $\denotesS{\Sigma}$'', the best way to understand is to consider we are defining a function $\denotesS{\goodType{\Gamma}{\cC \_}{}}$ as the interpretation for $\goodType{\Gamma}{\cC \_}{}$ that mapping $\denotesS{\sigma}$ to $\denotesS{\goodType{\Gamma}{\cC \sigma}{}}$ also defined inductively on the length of the signature. The case is similar for $\denotesS{\cL \_}$, $\denotesS{\cCt \_}$ and  $\denotesS{\Sigma \ \_ \ \_}$. \textbf{This is actually the formulation in Agda}---QIIT syntax consider $\cC : "Sig"~\Gamma~n \to "Ty"~\Gamma$ as a constructor and our model needs to interpret this constructor to a function $\cC^S : "Sig"^S \Gamma^S~n \to "Ty"^S~\Gamma^S$ with each judgement type also interpreted.

@@

\begin{figure}[!htb]
  \centering
\begin{minipage}{0.8\linewidth}
  \begin{minted}[mathescape, escapeinside=||]{Agda}
    data MW {i}{j} (σ : |$\denotesS{\texttt{WSig i }\cdot\texttt{ S}}$|) : Set (lsuc i) where
      MWsup : |$n$| < S → (a : |$\pi^n_1$| σ) → ((|$\pi^n_2$| σ) a → MW σ) 
              → MW σ
      \end{minted}
\end{minipage}
\caption{Modeling Inductive Type}\label{fig:model-ind-type}
\end{figure}
